Estimating Information Transfer from Different Sources in High-dimensional Linear Regression

### 2021-07-22 revision_journal/abstract.tex
intro wip

\iffalse
\begin{frontmatter}

\title{Precise Asymptotics of Transferring from Different Regression in High Dimensions}

\runtitle{Precise Asymptotics of Transferring from Different Regression}

\begin{aug}
	\author[A]{\fnms{Fan} \snm{Yang}\ead[label=e1]{}},
	\author[B]{\fnms{Hongyang R.} \snm{Zhang}\ead[label=e2]{}},
	\author[C]{\fnms{Sen} \snm{Wu}\ead[label=e3]{}},\\
	\author[A]{\fnms{Weijie J.} \snm{Su}\ead[label=e4]{}},
	\and
	\author[C]{\fnms{Christopher} \snm{R\'e}\ead[label=e5]{}}
	\address{${ }^{1}$University of Pennsylvania, ${ }^{2}$Northeastern University, ${ }^{3}$Stanford University
}
\end{aug}
\fi

\begin{abstract}
	Hard parameter sharing for multi-task learning is widely used in empirical research despite the fact that its generalization properties have not been well established in many cases. This paper studies its generalization properties in a fundamental setting: How does hard parameter sharing work given multiple linear regression tasks? We develop new techniques and establish a number of new results in the high-dimensional setting, where the sample size and feature dimension increase at a fixed ratio. First, we show a sharp bias-variance decomposition of hard parameter sharing, given multiple tasks with the same features. Second, we characterize the asymptotic bias-variance limit for two tasks, even when they have arbitrarily different sample size ratios and covariate shifts. We also demonstrate that these limiting estimates for the empirical loss are incredibly accurate in moderate dimensions. Finally, we explain an intriguing phenomenon where increasing one task's sample size helps another task initially by reducing variance but hurts eventually due to increasing bias. This suggests progressively adding data for optimizing hard parameter sharing, and we validate its efficiency in text classification tasks.
\end{abstract}

%\begin{keyword}[class=MSC2020]
%\kwd[Primary ]{62J05}
%\kwd{60B20}
%\kwd[; secondary ]{62E20, 62H10}
%\end{keyword}

%\begin{keyword}
%\kwd{Transfer learning, random matrix theory, covariate shift, sample covariance matrices.}
%\end{keyword}

%\end{frontmatter}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please use \tableofcontents for articles %%
%% with 50 pages and more                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\tableofcontents

>>>>

\begin{abstract}
%	Hard parameter sharing for multi-task learning is widely used in empirical research despite the fact that its generalization properties have not been well established in many cases. This paper studies its generalization properties in a fundamental setting: How does hard parameter sharing work given multiple linear regression tasks? We develop new techniques and establish a number of new results in the high-dimensional setting, where the sample size and feature dimension increase at a fixed ratio. First, we show a sharp bias-variance decomposition of hard parameter sharing, given multiple tasks with the same features. Second, we characterize the asymptotic bias-variance limit for two tasks, even when they have arbitrarily different sample size ratios and covariate shifts. We also demonstrate that these limiting estimates for the empirical loss are incredibly accurate in moderate dimensions. Finally, we explain an intriguing phenomenon where increasing one task's sample size helps another task initially by reducing variance but hurts eventually due to increasing bias. This suggests progressively adding data for optimizing hard parameter sharing, and we validate its efficiency in text classification tasks.
\end{abstract}

### 2021-07-22 revision_journal/intro.tex
intro wip

Many modern machine learning methods
%Hard parameter sharing (HPS) for multi-task learning is widely used in empirical research and goes back to the seminal work of \cite{C97}.
Recent work has revived interests in this approach because it improves performance and reduces the cost of collecting labeled data \cite{R17}.
It is generally applied by sharing the feature layers between all tasks while keeping an output layer for every task.
Often, hard parameter sharing offers two critical advantages if successfully applied.
First, it reduces model parameters since all tasks use the same feature space.
Second, it reduces the amount of labeled data needed from each task by augmenting the entire training dataset.

Hard parameter sharing works as an inductive transfer mechanism and a regularizer that reduces overfitting, both of which have great intuitive appeal \cite{R17}.
For example, by restricting the shared space's size, HPS encourages information sharing among multiple tasks \cite{KD12}.
Another source of inductive bias comes from the tasks and depends on datasets' properties such as sample sizes and task covariances \cite{WZR20}.
However, how these dataset properties impact HPS has not been well established.
%It becomes increasingly important to understand HPS' formal generalization properties.
Part of the challenge may be that HPS' generalization performance depends intricately on the sample size ratios and covariate shifts between tasks, and is not amenable to standard concentration results.
Previous results based on Rademacher complexity or VC dimensions have considered cases where all tasks' sample sizes are equal to logarithmic factors of the feature dimension \cite{B00,MPR16}, and when all tasks' sample sizes increase simultaneously \cite{AZ05,M06}.
%For, the generalization error scales down as the sample sizes of all tasks increase, when applied to the multi-task setting \cite{B00,AZ05,M06,MPR16,WZR20}.

\begin{figure*}[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.745\textwidth,valign=t]{figures/sample_ratio_c2_400.eps}
		\caption{Covariate shift}
		\label{fig_intro_arch}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.745\textwidth,valign=t]{figures/sample_ratio_c2_400.eps}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size_b}
	\end{subfigure}
	\vspace{-0.1in}
	\caption{
	Left: an illustrative picture of HPS.
	Right: an illustrative example of using HPS for two tasks $X_1, Y_1$ and $X_2, Y_2$ with sample size $n_1, n_2$, respectively.
	Increasing $n_1/n_2$ decreases task two's prediction loss initially but increase afterward. This phenomenon occurs due to different bias-variance tradeoffs as $n_1/n_2$ increases. Our result provides an estimated loss (solid line) that accurately matches the empirical loss (dotted line).
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
\end{figure*}

This paper presents new techniques to study hard parameter sharing and establishes a number of new results.
We consider regression analysis, which is arguably one of the most fundamental problems in statistics and machine learning.
We are interested in the \textit{high-dimensional} setting, where each dataset's sample size and feature dimension grow linearly instead of logarithmically.
This setting captures the fact that a single task's sample size is usually insufficient for accurate learning in many applications.
For example, if a dataset's sample size is only a constant factor of dimension in linear regression, the variance is also constant (cf. Fact \ref{fact_tr}).
The high-dimensional setting is challenging but is crucial for understanding how datasets' sample sizes impact generalization performance.

\subsection{Summary of results}
\todo{mention connection to pooling \cite{hanneke2020no}}
%\subsection{Summary of Results}

@@

Our second result (Theorem \ref{thm_main_RMT}) applies to two tasks with arbitrarily different sample size ratios and covariate shifts.
While we can no longer characterize $f(A, B)$'s global minimum because of non-convexity, we can still provide a sharp bias-variance tradeoff of any local minimizer's prediction loss for both tasks.
Despite being a simple setting, we observe several non-trivial phenomena by varying sample size ratios and covariate shifts between the two tasks.
See Figure \ref{fig_intro_sample_size_b} for an illustration of the former.
Consequently, using our precise loss estimates, we observe several qualitative properties of HPS for varying dataset properties.
\begin{itemize}
	\item \textit{Sample efficiency (Example \ref{ex_same_cov})}:
	One advantage of combining multiple datasets is that the requirement for labeled data reduces compared to single-task learning, a phenomenon that \cite{ZSSGM18} has observed empirically.
	Our results further imply that HPS's sample efficiency depends on model-specific variances across tasks vs. the noise variance and is generally high when the latter is large.

@@

	\item \textit{Covariate shift (Example \ref{ex_covshift})}: In addition to sample sizes, variance also scales with two datasets' covariate shifts. For a large sample size ratio, HPS's  variance is smallest when there is no covariate shift. Counterintuitively, for a small sample size ratio, having covariate shifts reduces variance through a complementary spectrum. We achieve this result through a novel characterization on the inverse of the sum of two sample covariance matrices with arbitrary covariate shifts. See our discussion of proof techniques below for details.
\end{itemize}

@@

%\section{Predication loss for HPS estimator}\label{sec_HPS_loss}

@@

%\subsection{Notations}

%for a small constant $\tau>0$.

\subsection{Organizations}
The rest of this paper is organized as follows.
In Section \ref{sec_same}, we present the bias-variance decomposition for hard parameter sharing.
In Section \ref{sec_diff}, we describe how varying sample sizes and covariate shifts impact hard parameter sharing using random matrix theory.
In Sections \ref{sec_simulation}, we validate our results in simulations.
In Section \ref{sec_conclude}, we summarize our work and discuss future work.
Section \ref{sec_text} describes our study on text classification tasks.
Section \ref{app_proof_error_same_cov}, \ref{appendix RMT}, and \ref{app_iso_cov} present proofs of our results.

>>>>

%Many modern machine learning methods
%Hard parameter sharing (HPS) for multi-task learning is widely used in empirical research and goes back to the seminal work of \cite{C97}.
%Recent work has revived interests in this approach because it improves performance and reduces the cost of collecting labeled data \cite{R17}.
%It is generally applied by sharing the feature layers between all tasks while keeping an output layer for every task.
%Often, hard parameter sharing offers two critical advantages if successfully applied.
%First, it reduces model parameters since all tasks use the same feature space.
%Second, it reduces the amount of labeled data needed from each task by augmenting the entire training dataset.

%Hard parameter sharing works as an inductive transfer mechanism and a regularizer that reduces overfitting, both of which have great intuitive appeal \cite{R17}.
%For example, by restricting the shared space's size, HPS encourages information sharing among multiple tasks \cite{KD12}.
%Another source of inductive bias comes from the tasks and depends on datasets' properties such as sample sizes and task covariances \cite{WZR20}.
%However, how these dataset properties impact HPS has not been well established.
%It becomes increasingly important to understand HPS' formal generalization properties.
%Part of the challenge may be that HPS' generalization performance depends intricately on the sample size ratios and covariate shifts between tasks, and is not amenable to standard concentration results.
%Previous results based on Rademacher complexity or VC dimensions have considered cases where all tasks' sample sizes are equal to logarithmic factors of the feature dimension \cite{B00,MPR16}, and when all tasks' sample sizes increase simultaneously \cite{AZ05,M06}.
%For, the generalization error scales down as the sample sizes of all tasks increase, when applied to the multi-task setting \cite{B00,AZ05,M06,MPR16,WZR20}.

%This paper presents new techniques to study hard parameter sharing and establishes a number of new results.
%We consider regression analysis, which is arguably one of the most fundamental problems in statistics and machine learning.
%We are interested in the \textit{high-dimensional} setting, where each dataset's sample size and feature dimension grow linearly instead of logarithmically.
%This setting captures the fact that a single task's sample size is usually insufficient for accurate learning in many applications.
%For example, if a dataset's sample size is only a constant factor of dimension in linear regression, the variance is also constant (cf. Fact \ref{fact_tr}).
%The high-dimensional setting is challenging but is crucial for understanding how datasets' sample sizes impact generalization performance.

\subsection{Setup and motivating examples}\label{sec_HPS_loss}

\subsection{Summary of results}
\todo{mention connection to pooling \cite{hanneke2020no}}

@@

Our second result (Theorem \ref{thm_main_RMT}) applies to two tasks with arbitrarily different sample size ratios and covariate shifts.
While we can no longer characterize $f(A, B)$'s global minimum because of non-convexity, we can still provide a sharp bias-variance tradeoff of any local minimizer's prediction loss for both tasks.
Despite being a simple setting, we observe several non-trivial phenomena by varying sample size ratios and covariate shifts between the two tasks.
%See Figure \ref{fig_intro_sample_size_b} for an illustration of the former.
Consequently, using our precise loss estimates, we observe several qualitative properties of HPS for varying dataset properties.
%\begin{itemize}
%	\item \textit{Sample efficiency (Example \ref{ex_same_cov})}:
%	One advantage of combining multiple datasets is that the requirement for labeled data reduces compared to single-task learning, a phenomenon that \cite{ZSSGM18} has observed empirically.
%	Our results further imply that HPS's sample efficiency depends on model-specific variances across tasks vs. the noise variance and is generally high when the latter is large.

@@

	\item \textit{Covariate shift (Example \ref{ex_covshift})}: In addition to sample sizes, variance also scales with two datasets' covariate shifts. For a large sample size ratio, HPS's  variance is smallest when there is no covariate shift. Counterintuitively, for a small sample size ratio, having covariate shifts reduces variance through a complementary spectrum. We achieve this result through a novel characterization on the inverse of the sum of two sample covariance matrices with arbitrary covariate shifts. See our discussion of proof techniques below for details.
%\end{itemize}

@@

@@

\subsection{Organizations}
The rest of this paper is organized as follows.
In Section \ref{sec_HPS}, we formally define the data model and the assumptions.
We describe a bias-variance decomposition of the HPS estimator, and connect the decomposition to random matrix theory.
In Section \ref{sec_main}, we show precise estimates for the excess risk of the HPS estimator under covariate and model shifts.
In Section \ref{sec_same}, we extend our setup to transferring from multiple data sources under model shift.
In Section \ref{sec_exp}, we evaluate the performance of HPS estimators against several commonly used transfer learning estimators, and present two implications for migating covariate and model shift in neural networks.
Section \ref{sec_conclude} concludes the paper and discusses potential future directions.
Section \ref{app_proof_error_same_cov}, \ref{appendix RMT}, and \ref{app_iso_cov} present proofs of our results.

### 2021-07-22 revision_journal/main_arxiv.tex
reorg wip
update intro wip
sec 2 wip
Updates from Overleaf
finish sec 4 edits
finish sec 3.2
start intro
intro wip

\def\shownotes{1}

\usepackage{natbib}
\setcitestyle{authoryear,open={(},close={)}}
\renewcommand{\cite}{\citep}
\newcommand{\yell}[1]{{\color{red}#1}}
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
\usepackage{authblk}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}

@@

	\iffalse
	\vspace{0.15in}
	\author[1]{Hongyang R. Zhang\thanks{HZ and FY contributed equally to the theoretical results. SW conducted the experimental analysis. Correspondence to hrzhang@northeastern.edu or fyang75@wharton.upenn.edu.}}
	\author[2]{Fan Yang\samethanks}
	\author[3]{Sen Wu\samethanks}
	\author[2]{Weijie J. Su}
	\author[3]{Christopher R\'e}
	\vspace{0.15in}
	\affil[1]{Khoury College of Computer Sciences, Northeastern University}
	\affil[2]{Department of Statistics, The Wharton School, University of Pennsylvania}
	\affil[3]{Department of Computer Science, Stanford University}
	\fi
	\title{Precise Asymptotics of Transferring from Different Regression in High Dimensions}
	\maketitle
	\input{abstract}
	\input{intro}
	\input{results}
	\input{examples}
	\input{multitask}
	\input{exp}
	\input{conclusions}
	\bibliographystyle{plainnat}
	\bibliography{rf,ref_mtl}

>>>>

\def\shownotes{1}
\setlength{\columnsep}{0.35in}
\usepackage{natbib}
\setcitestyle{authoryear,open={(},close={)}}
\renewcommand{\cite}{\citep}
\newcommand{\yell}[1]{{\color{red}#1}}
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
%\usepackage{authblk}
\usepackage[margin=1in]{geometry}
%\usepackage{hyperref}

@@

	\title{Estimating Information Transfer from Different Sources in High-dimensional Linear Regression}

    \author{Fan Yang\thanks{University of Pennsylvania. fyang75@wharton.upenn.edu}\and Hongyang R. Zhang \thanks{Northeastern University. ho.zhang@northeastern.edu. Part of this work was done while the author was at UPenn.}\and Sen Wu\thanks{Stanford University. senwu@stanford.edu}\and Weijie J. Su \thanks{University of Pennsylvania. suw@wharton.upenn.edu}\and Christopher R\'e\thanks{Stanford University. chrismre@stanford.edu}}

	\maketitle
	\input{abstract}
	\input{intro}
	\input{setup}
	\input{results}
	\input{exp}
    \input{multitask.tex}
	\input{conclusions}
	\bibliographystyle{plainnat}
	\bibliography{rf,rf_tl,ref_mtl}
